\section{Summary}
In recent years, quantum computing has undergone significant developments and has established its supremacy in many application domains.
Unfortunately, modern quantum computing is still positioned in the Noisy Intermediate-Scale Quantum (NISQ) era that is limited by the number of qubits, short qubit lifetime, and imperfect operations.
While quantum hardware is accessible to the public through the cloud environment, a robust and efficient quantum circuit simulator is necessary to investigate the constraints and foster quantum computer development, such as quantum algorithm development and quantum device architecture exploration.
In this paper, we observe that most of the publicly available quantum circuit simulators (e.g., QISKit from IBM) are not optimized and suffer from slow simulation and poor scalability.
To this end, we systematically studied the deficiencies in modern quantum simulators and propose Q-GPU, a framework that leverages GPUs with comprehensive optimizations to allow efficient and scalable quantum circuit simulation (QCS).
Specifically, Q-GPU features i) proactive state amplitude transfer, ii) zero state amplitudes pruning, iii) delayed qubit involvement, and iv) non-zero state compression.
Experimental results across eight representative quantum circuits indicate that Q-GPU significantly improves the simulation performance over the state-of-the-art GPU-based QCS by 2.53$\times$ on average.
It also outperforms the most recent OpenMP CPU implementation, the Google Qsim-Cirq simulator, and the Microsoft QDK simulator.
