This thesis is first dedicated to my family -- my mother Kailas, my father Navnit, and my sisters Sri and Krupa.
I have always found support and encouragement from my parents, and my efforts through graduate school were made possible through this energy.
I have looked up to my older sister, Sri, for her dedicated pursuit of her interests both professionally and personally.
Krupa, my younger sister, was also an inspiration as she was always empathetic and engaged in interpersonal communication and collectively with her surrounding community as a whole.
She will be remembered fondly.

Kenneth D.~Jordan was an excellent mentor and provided support and guidance through my entire graduate career.
Ken has shaped the way I approach problems as he had an uncanny ability to cut to the heart of an issue and find a simpler system that displays the same characteristics.
Former and current graduate students in the Jordan group have been extremely helpful, including but not limited to Amanda Dumi, Kevin Gasperich, Arailym Kairalapova, and Devin Mulvey.
There are countless times where I have turned to my peers for advice and found invaluable advice.

I would also like to thank the other theory professors at the University of Pittsburgh, Dr.~Chong, Dr.~Hutchison, Dr.~Liu, Dr.~Coalson, and formerly Dr.~Lambrecht.
I have also benefited from conversations with other theory group graduate students including Eric Berquist, Keith Werling, Dakota Folmsbee, Danielle Hiener, Anthony Bogetti, and Sanjeev Gautam.

Anouar Benali was extremely influential to my development and provided a lot of scientific and financial support.
I'd like to thank a number of other collaborators and external friends: Brenda Rubenstein, Hongxia Hao, James Shee, Tina Mihm, Paul Kent, Ye Luo, and Hyeondeok Shin.

I would not have progressed to this point in my scientific career without a lot of guidance along the way.
From my undergraduate days at Washington and Jefferson college, interactions with Dr. Iuliucci and Dr. McCracken led me to enjoy science and eventually theoretical chemistry.
I began my graduate studies at Duquesne University under Dr.~Jeffry D.~Madura, who taught me to be inquisitive, work hard, and to have fun while doing science.
His memory will live on in the work of his students.
I have to thank the Madura group members who have all helped me along the way particularly Mike Wasko, Matt Srnec, Emily Benner, Kendy Guarinoni, Riley Workman, and Amanda Dumi.

Finally, I want to thank my friends.
Amanda Dumi -- through many years of graduate school, we became best friends.
From our start at Duquesne in the same year, we had a long, and sometimes difficult, path through graduate school.
I enjoyed the numerous hours spent together working on problem sets, and later, collaborations.
Justin Kinslow, Patrick Brady, Sean Rieger, and Andrew Giel -- my close friends of many years.
I appreciate your patience as I often flaked on plans through the years, and I look forward to many trips in the future.
Rajan Plumley, Riley Workman, Sean Gracy, Nick Gilbert, and Evan Appod -- I appreciate the many hours spent virtually to decompress from the stresses of life.
